Generating Performance Improvement Suggestions by using Cross-Organizational Process Mining

### 2015-11-21 4_methodology/methodology.tex
no message

\subsection{Performance Indicator Analysis}
\label{subsec:performance-indicator-analysis}
Performance indicator analysis stage focuses on calculating and analyzing the performance values using the event logs and mined process models. This stage consists of mainly two steps as 
\begin{inparaenum}[\itshape a\upshape)]
\item alignment and calculation of performance indicators; and
\item clustering of organizations based on their performance values.
\end{inparaenum}
In order to evaluate the performance of an organization based on their process models and past activities; there is a number of indicators in time dimension, cost dimension and utilization \cite{van2011process}. However, in this study, process related performance values are considered since differences in the process models are studied in the next stages. To this aim, the following performance indicators are calculated and analyzed in this stage of methodology:
\begin{description}
	\item[Average Time Between Activities] For each activity in the process model, average time to reach the next activity is calculated. This is a simple but powerful performance metric for organizations since it can yield the average time to complete one task based on a starting point. From the performance perspective, organizations want to minimize average time between activities to increase their throughput \cite{van2012replaying}. This notion can be defined as follows:
  \theoremstyle{definition}
  \begin{definition}{}
  Average time between activity $A$ and $B$ in organization $i$ is $AvgTime_{A\rightarrow B}^{i} = \frac{\sum_{Case\ c \in Event Log_{i}} TimeBetween_{c}(A,B)}{|Occurences_{Event\ Log_{i}}(A,B)|}$ where
    \begin{enumerate}
      \item $TimeBetween_{c}(A, B) = EndTime_{c}(B) - StartTime_{c}(A)$
      \item $StartTime_{c}(A)$ is start time of activity $A$ in case $c$,
      \item $EndTime_{c}(B)$ is end time of activity $B$ in case $c$,
      \item $|Occurences_{Event Log_{i}}(A, B)|$ is number of occurrences of activity $A$ followed by $B$ in  $Event\ Log_{i}$.
    \end{enumerate}
  \end{definition}
	\item[Standard Deviation of Time Between Activities] Time between activities in real life is not stable and they deviate due to various reasons such as the user responsible of tasks, size and the content of tasks or seasonality \cite{van2011process}. On the other hand, organizations want to be confident about their processes and therefore they want to minimize the deviation in the time between activities. Minimized deviation in time helps organizations to plan, act and re-organize the activities in the processes with high accuracy \cite{van2012replaying}. With the same approach above, the following formulation can be defined:
  \theoremstyle{definition}
  \begin{definition}{}
  Standard deviation time between activity $A$ and $B$ in organization $i$ is $StdDevTime_{A\rightarrow B}^{i} = \sqrt{\frac{\sum_{Case\ c \in Event Log_{i}} [TimeBetween_{c}(A, B) - AvgTime_{A\rightarrow B}^{i}]^{2}}{Occurences_{Event\ Log_{i}}(A,B)} }$ 
  \end{definition}
\end{description}

@@

\subsection{Mismatch Pattern Analysis}
\label{subsec:mismatch-pattern-analysis}
In order to learn from other organizations, it is necessary to spot the differences between process models of different organizations. In this phase, differences between process models will be revealed by the mismatch patterns which are defined by Dijkman \cite{dijkman2007mismatch}. Since performance indicators are calculated based on a starting and ending point in the process model, the same approach is applied to locate mismatch patterns. In other words, differences of process models are located through a starting activity to an ending activity. With this aim, each mismatch pattern and its analyzers are defined by extending the following definitions. For each organization, mismatch pattern analyzers are pipelined and mismatch patterns are stored for further analysis.
\theoremstyle{definition}
    \begin{definition}
    Mismatch Pattern is a tuple ${Mismatch\ Patten} = (O_{1}, O_{2}, ExtensionData, A_{start}, A_{end}) $ where 
    \begin{enumerate}
      \item $O_{1}$ is the first organization and \item $O_{2}$ is the second organization in between the pattern occurs,
      \item $ExtensionData$ is a set of tuples where mismatch related information is recorded, 
      \item $A_{start}$ and $A_{end}$ are starting and ending points to check mismatch patterns.
    \end{enumerate}
    \end{definition}
\theoremstyle{definition}

\begin{definition}
    Mismatch Pattern Analyzer is a function $MismatchPatternAnalyzer(O_{1}, O_{2}, A_{start}, A_{end})$ and it returns a set of \textit{Mismatch Pattern} for the organization $O_{1}$ compared to $O_{2}$ for the activities between $A_{start}$ and $A_{end}$.
\end{definition}

@@

Algorithm of recommendation generation function is based on the idea of checking other clusters for a significant change in performance indicators, where significance is defined by the threshold provided by user. After finding significant difference, all organizations in other clusters are checked against mismatch patterns with the starting and ending activities defined in performance indicators. Only mismatches which are located between the activities that causes high level of difference in performance indicators are analyzed. This approach is formalized in Algorithm~\ref{algo:recgen}.
 \begin{algorithm}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead 
\KwIn{$O$ organization, $C$ Cluster Analysis Data, $P$ performance difference threshold}
\KwOut{$Recommendations$ a set of recommendations}
$Recommendations \leftarrow \{\}$ \;
$i \leftarrow C(Assignments(O))$ \;
\For{$Centroid \in C(Cluster Centroids_{i})$} { 
  \For{$Centroid' \in C(Cluster Centroids_{j})\ i\neq j$} { 
    \If{$Centroid(A_{start}) = Centroid'(A_{start})\ \&\ Centroid(A_{end}) = Centroid'(A_{end})$} {
      \If{ ($\left |  Centroid(Value) -  Centroid'(Value)\right | \div Centroid(Value)) \geq P$  } {
        $A_{start} \leftarrow Centroid(A_{start})$ \;
        $A_{end} \leftarrow Centroid(A_{end})$ \;
        $MismatchPatterns \leftarrow \{\}$ \;
        \For{$O' \in C(Assignments(j))$} {
          $MismatchPatterns$ \leftarrow  Mismatch Pattern Analysis($O$,$O'$,$A_{start}$,$A_{end}$) \;
        }
        $Recommendations$ \leftarrow  Recommendation($O$,$A_{start}$,$A_{end}$, $MismatchPatterns$) \;
      }
    }
  }
}
\Return{$Recommendations$} \;
\caption{Recommendation Generation}
\label{algo:recgen}
\end{algorithm}

>>>>

\subsection{Performance Indicator Analysis}
\label{subsec:performance-indicator-analysis}
Performance indicator analysis stage focuses on calculating and analyzing the performance values using the event logs and mined process models. This stage consists of mainly two steps as 
\begin{inparaenum}[\itshape a\upshape)]
\item alignment and calculation of performance indicators; and
\item clustering of organizations based on their performance values.
\end{inparaenum}
In order to evaluate the performance of an organization based on their process models and past activities; there is a number of indicators in time dimension, cost dimension and utilization \cite{van2011process}. However, in this study, process related performance values are considered since differences in the process models are studied in the next stages. To this aim, the following performance indicators are calculated:
\begin{description}
	\item[Average Time Between Activities] This is a simple but powerful performance metric for organizations since it can yield the average time to complete one task based on a starting point. From the performance perspective, organizations want to minimize average time between activities to increase their throughput \cite{van2012replaying}. This notion can be defined as follows:
  \theoremstyle{definition}
  \begin{definition}{}
  Average time between activity $A$ and $B$ in organization $i$ is $AvgTime_{A\rightarrow B}^{i} = \frac{\sum_{Case\ c \in Event Log_{i}} TimeBetween_{c}(A,B)}{|Occurences_{Event\ Log_{i}}(A,B)|}$ where
    \begin{enumerate}
      \item $TimeBetween_{c}(A, B) = EndTime_{c}(B) - StartTime_{c}(A)$
      \item $StartTime_{c}(A)$ is start time of activity $A$ in case $c$,
      \item $EndTime_{c}(B)$ is end time of activity $B$ in case $c$,
      \item $|Occurences_{Event Log_{i}}(A, B)|$ is number of occurrences of activity $A$ followed by $B$ in  $Event\ Log_{i}$.
    \end{enumerate}
  \end{definition}
	\item[Standard Deviation of Time Between Activities] Time between activities in real life is not stable and they deviate due to various reasons such as the user responsible of tasks, size and the content of tasks or seasonality \cite{van2011process}. On the other hand, organizations want to be confident about their processes and therefore they want to minimize the deviation in the time between activities. Minimized deviation in time helps organizations to plan, act and re-organize the activities in the processes with high accuracy \cite{van2012replaying}. With the same approach above, the following formulation can be defined:
  \theoremstyle{definition}
  \begin{definition}{}
  Standard deviation time between activity $A$ and $B$ in organization $i$ is $StdDevTime_{A\rightarrow B}^{i} = \sqrt{\frac{\sum_{Case\ c \in Event Log_{i}} [TimeBetween_{c}(A, B) - AvgTime_{A\rightarrow B}^{i}]^{2}}{Occurences_{Event\ Log_{i}}(A,B)} }$ 
  \end{definition}
\end{description}

@@

\subsection{Mismatch Pattern Analysis}
\label{subsec:mismatch-pattern-analysis}
In order to learn from other organizations, it is necessary to spot the differences between process models of different organizations. In this phase, differences between process models will be revealed by the mismatch patterns which are defined by Dijkman \cite{dijkman2007mismatch}. Since performance indicators are calculated based on a starting and ending point in the process model, the same approach is applied to locate mismatch patterns. In other words, differences of process models are located through a starting activity to an ending activity. With this aim, each mismatch pattern and its analyzers are defined by extending the following definitions. For each organization, mismatch pattern analyzers are pipelined and mismatch patterns are stored for further analysis.
\theoremstyle{definition}
    \begin{definition}
    Mismatch Pattern is a tuple ${Mismatch\ Patten} = (O_{1}, O_{2}, ExtensionData, A_{start}, A_{end}) $ where 
    \begin{enumerate}
      \item $O_{1}$ is the first organization and \item $O_{2}$ is the second organization in between the pattern occurs,
      \item $ExtensionData$ is a set of tuples where mismatch related information is recorded, 
      \item $A_{start}$ and $A_{end}$ are starting and ending points to check mismatch patterns.
    \end{enumerate}
    \end{definition}
\theoremstyle{definition}
\begin{definition}
    Mismatch Pattern Analyzer is a function $MismatchPatternAnalyzer(O_{1}, O_{2}, A_{start}, A_{end})$ and it returns a set of \textit{Mismatch Pattern} for the organization $O_{1}$ compared to $O_{2}$ for the activities between $A_{start}$ and $A_{end}$.
\end{definition}

@@

Algorithm of recommendation generation function is based on the idea of checking other clusters for a significant change in performance indicators, where significance is defined by the threshold provided by user. Only mismatches which are located between the activities that causes high level of difference in performance indicators are analyzed. This approach is formalized in Algorithm~\ref{algo:recgen}.
 \begin{algorithm}
\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead 
\KwIn{$O$ organization, $C$ Cluster Analysis Data, $P$ performance difference threshold}
\KwOut{$Recommendations$ a set of recommendations}
$Recommendations \leftarrow \{\}$ \;
$i \leftarrow C(Assignments(O))$ \;
\For{$Centroid \in C(Cluster Centroids_{i})$} { 
  \For{$Centroid' \in C(Cluster Centroids_{j})\ i\neq j$} { 
    \If{$Centroid(A_{start}) = Centroid'(A_{start})\ \&\ Centroid(A_{end}) = Centroid'(A_{end})$} {
      \If{ ($\left |  Centroid(Value) -  Centroid'(Value)\right | \div Centroid(Value)) \geq P$  } {
        $A_{start} \leftarrow Centroid(A_{start})$ \;
        $A_{end} \leftarrow Centroid(A_{end})$ \;
        $MismatchPatterns \leftarrow \{\}$ \;
        \For{$O' \in C(Assignments(j))$} {
          $MismatchPatterns$ \leftarrow  Mismatch Pattern Analysis($O$,$O'$,$A_{start}$,$A_{end}$) \;
        }
        $Recommendations$ \leftarrow  Recommendation($O$,$A_{start}$,$A_{end}$, $MismatchPatterns$) \;
      }
    }
  }
}
\Return{$Recommendations$} \;
\caption{Recommendation Generation}
\label{algo:recgen}
\end{algorithm}
